Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results

### 2014-10-30 workshop-paper/paper.tex
Hectors last version

\section{Related and Futue Work}

- Relation  to Novelty-based search (Stanley)
- IW probes vs UCT probes; UCT using novelty with noise in rollout???
- Structured Exploration of  Factored Spaces
- Dijkstra postprocessing, properties for optimization
- Learning in same instance, across diff instances of same game, etc.
- Learning controllers vs. on-line planning, etc

>>>>

\section{Exploration and Exploitation: Discussion}

The notion of width underlying the iterated width algorithm was developed in the context
of classical planning in order to understand why most of the hundreds of existing benchmarks
resulting from planning competitions appear to be relatively simple for current planners, even
though classical planning is PSPACE-complete \cite{bylander}. The partial answer is that 
most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
when goals contain a single atom. At the same time, most problems with multiple atomic goals are 
also easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.

In the Iterated Width algorithm, the key notion is the \emph{novelty measure} of a state in the underlying
breadth-first search. In particular, a state gets a novelty 1 measure in the search if it makes true,
some atom that up to that point had been false in all previously generated states. In the procedure IW(1)
all other states are pruned right away, as if they were ``duplicate'' states. The algorithms IW and IW(k)
are pure  exploration algorithms that make no use at all of the information about the goal, if there is a goal,
or about the  reward function to optimize otherwise. They are  thus  pure exploration algorithms guided by the novelty
measures, which use the structure of the states to provide a  structure to the exploration of the state space,
which in the case of IW is guaranteed to be complete.

The use of ``novelty measures'' for guiding an optimization search while ignoring the function that is being
optimized, appears to be  common to novelty-based search methods developed independently 
in the context of genetic algorithms \cite{lehman:novelty}. %%% bibtex for this below
In these methods individuals in the population are not ranked according to optimization function but in terms of how 
``novel'' thay are in relation to the rest of the population, thus encouraging diversity
and exploration, rather than (greedy) exploitation. The actual definition of novelty
in such a case is domain-dependent; for example, in the evolution of a controller
for guiding a robot in a maze, an individual controller  will not be ranked by how
close it takes the robot to the goal (the greedy measure), but how far it takes
the robot in relation to the locations that are reached by following the other controllers
(a diversity measure). The novelty measure used by IW, on the other hand, 
is domain-independent and it follows from the structure of the state
as captured by the problem variables.

% @article{lehman:novelty,
%  title={Abandoning objectives: Evolution through the search for novelty alone},
%  author={Lehman, Joel and Stanley, Kenneth O},
%  journal={Evolutionary computation},
%  volume={19},
%  number={2},
%  pages={189--223},
%  year={2011},
%  publisher={MIT Press}
%}

The balance between exploration and exploitation has received considerable attention in
reinforcement learning \cite{sutton:book}, where both are required for converging to an
optimal behavior. In the Atari games, as in other deterministic problems, however, ``exploration'' 
is not needed for optimality purposes, but just for improving the effectiveness of the 
local lookahead searches. Indeed, a best-first search algorithm guided by (discounted)
accumulated reward will deliver eventually  best moves, but it will not be as effective
over small time windows, where like breadth-first search it's likely not to find any rewards
at all. The UCT algorithm provides a method for balancing exploration and exploitation,
which can thus be effective over small time windows. The 2BFS algorithm above with two queues
that alternate, one guided by the novelty measures, and the other by the accumulated reward,
provides a different scheme. The first converges to the optimal behavior asymptotically; the 
second in a bounded number of steps. Alternately, novelty-guided probes, possibly with 
some noise added, could potentially pay off as  a  base policy for UCT in replacement 
of random rollouts, as they manage apparently to exploit the structure of many problems.

\section{Summary and On-going Work}

We have shown experimentally that width-based algorithms like IW(1) and 2BFS
that originate in work in classical planning, can be used in the Atari games
with a performance that approaches the state-of-the-art as captured by UCT 
planning, sometimes using a fraction of the time. We are currently addressing three issues, 
which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT
is often taking an order-of-magnitude more time per decision as 2BFS, so the 
comparison in not fair. Third, we need to revise the way duplicate states are handled 
in IW and 2BFS to ensure that the best paths among the paths defined by the states that have been expanded
are returned. This change is important from a theoretical point of view; it's not clear
what it's practical impact will be.

%%% - Learning in same instance, across diff instances of same game, etc.
%%%% - Learning controllers vs. on-line planning, etc
